### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization rate: second commit

---
title: "Assessing the Rates of Crimes Against the Person in Toronto "
author: Sofia Sellitto
date: February 3, 2023
date-format: long
abstract: Abstract 
format: pdf
bibliography: bibliography.bib
---

#

This data was free for use, courtesy of @opendatatoronto. We are using @R for analyzing this data.

## Loading Packages and Data

We import tidyverse by @tidyverse for convenience because it contains many other packages which will be useful for plotting and data manipulations. We also use @janitor for cleaning and @dplyr for manipulating data.

```{r}
## loading appropriate packages

library(tidyverse)
library(ggplot2)
library(knitr)
library(janitor)
library(dplyr)

```

```{r}
## Calling on cleaned data 
cleaned_data_dropped <- read_csv(here::here("inputs/data/cleaned_data.csv"))
head(cleaned_data_dropped)
```

```{r}

## summing count of crimes and making them its own group

cleaned_data_dropped_grouped <- cleaned_data_dropped |>
  group_by(subtype, reported_year) |>
  summarise(across(count, sum))

cleaned_data_dropped_grouped
```

```{r}
### making a data frame that includes the variable sex so that I can call on it for a graph

sex_data <- cleaned_data_dropped |>
  group_by(subtype,sex, reported_year) |>
  summarise(across(count, sum))

sex_data

```

```{r}

## calculating the mean number of crimes for each subcategory across 2014-2021

mean_data <- cleaned_data_dropped_grouped |>
    summarise(mean_count = mean(count))

mean_data

```

```{r}

## Creating a table to visually to display the mean number of crimes for each subcategory across 2014-2021

knitr::kable(mean_data,"simple", caption = "Average Number of victims of crimes in Toronto across 2014-2021, grouped by category of crime", 
             col.names=c("Category of Crime", "Average Number of Victims"), 
             align=c('l', 'c'))

```

```{r}

knitr::kable(head(cleaned_data_dropped_grouped), "simple", caption = "The first 6 rows of the data frame used to represent the number of victims according to category of crime in Toronto Across 2014 - 2021",
             col.names=c("Sub-Category of Crime", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

```

```{r}

knitr::kable(head(sex_data), "simple", caption = "The first 6 rows of the data frame used to represent the gender of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

```

See @figure1

```{r}
## creating a line graph that includes the number of victims according to category of crime in Toronto across 2014-2021

#| fig-cap: "Number of victims according to category of crime in Toronto Across 2014 - 2021"
#| label: figure1

cleaned_data_dropped_grouped |>
  ggplot(aes(x = reported_year, y = count, colour = subtype))+
  geom_point(linewidth = 5, alpha = 0.3)+
  geom_line(linewidth = 1)+
  theme_minimal()+
  labs(
    title = "Victims of Crime in Toronto Across 2014 - 2021",
    x = "Year",  
    y = "Number of Victims", 
    colour = "Category of Crime" )

```

See @figure2

```{r}

## calling on my data frame to create a bar plot of the sex of the victims across time

#| fig-cap: "How the victims of crimes varries across sex in Toronto from 2014-2021"
#| label: figure2

ggplot(sex_data, aes
       (x =reported_year, 
        y =count, 
        fill =sex))+
            geom_bar(position="dodge", stat="identity")+
            facet_wrap(~subtype)+
            scale_fill_manual(values=c("#C0AFE2", "#ADD8E6"))+
            theme_minimal()+
            labs(
            title = "Victims of Crime in Toronto Categorized by Sex Across 2014 - 2021",
            x = "Year",  
            y = "Number of Victims",
            legend = "Sex")

```

## References